Long-range correlations and their event-scale dependence in high-multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-08-10 paper.tex
First actions after IRC's comments

%%% Put your own title + short title here:
\title{Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV}
\ShortTitle{Long-range correlations in pp collisions}   % appears on right page headers

@@

Long-range azimuthal correlations are measured in high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by ALICE at LHC. 
These results are obtained in a relative pseudorapidity range of $1.6 < |\eta| < 1.8$ and transverse momentum region of $1 < p_{\mathrm T} < 5$~GeV/$c$ for the trigger and associated particles.
In addition to the conventional long-range correlation studies, the results are further investigated in events, where jets with a harder fragmentation are present.
The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events. 
These correlations also persist in the events containing jets. Furthermore data show that the ridge yields tend to increase in events holding $\ptjet > 10$~GeV/$c$.
The results are compared to $\pythiashoving$ calculations.  
$\pythiashoving$ model can describe low $p_{\mathrm T}$  ridge yields but can't give good description of $p_{\mathrm T}$ dependence both for the unbiased and jet events.
The results open new aspects of studying the small collision systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

>>>>

%%% Put your own title + short title here:
\title{Long-range correlations and their event-scale dependence in high-multiplicity pp collisions at $\sqrt{s} = 13$~TeV}
\ShortTitle{Long-range correlations in pp collisions}   % appears on right page headers

@@

Long-range azimuthal correlations are measured in high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by ALICE at the LHC. The measurement was carried out for trigger and associated particles having transverse momenta in the range $1 < p_{\mathrm T} < 5$~GeV/$c$. Trigger and associated particles were separated by a large pseudorapidity interval $1.6 < |\Delta \eta |<1.8$.
%These results are obtained in a relative pseudorapidity range of $1.6 < |\eta| < 1.8$ and transverse momentum region of  for the trigger and associated particles.
In addition to the conventional long-range correlation studies, the results are further investigated in events with a high-pT leading track or jet.
%where jets with a harder fragmentation are present.
The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
%The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events.
Measured ridge yield is consistent with the results from the CMS experiments for unbiased high-multiplicity events.
These long-range correlations are also seen in the events containing jets. Furthermore, data show that the ridge yields tend to grow with transverse momentum of the jet or leading track. %increase in events holding $\ptjet > 10$~GeV/$c$.
The results are compared to $\pythiashoving$ calculations.  
$\pythiashoving$ model can describe low $p_{\mathrm T}$ ridge yields up to 1.5 GeV/$c$ but can not give good description of $p_{\mathrm T}$ dependence both for the unbiased and jet-tagged events.
The results open new aspects of studying the small collision systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

### 2020-08-10 sec_ana.tex
First actions after IRC's comments

%LHC at CERN produces various interesting phenomena including the ridge effect that is the main topic of the document in pp collisions with the highest center-of-mass energy in the world.
%Recent center-of-mass energy in pp collisions reaches up to $\sqrt{s} = 13$~TeV during the last LHC Run 2 period. 
The analysis is based on the data sets in pp collisions at $\sqrt{s} = 13$~TeV collected from 2016 to 2018 in the LHC Run 2 period. The full description of ALICE detector in the LHC Run 2 can be found in Refs.~\cite{Aamodt:2008zz,Abelev:2014ffa}. The present analysis mainly utilizes the V0~\cite{Abbas:2013taa}, ITS (Inner Tracking System)~\cite{aliceITS} and TPC (Time Projection Chamber)~\cite{aliceTPC} detectors.

The V0 detector consists of two rings, V0-A and V0-C, each made of 32 scintillator tiles, covering the full azimuthal angle within $2.8 < \eta < 5.1$ and $-3.7 < \eta < -1.7$, respectively. The V0 provides trigger and estimation of event multiplicity. A sample of events including higher numbers of produced particles is obtained with a high multiplicity trigger, which is achieved by requiring 5 times of mean signal amplitudes of energy deposit by charged particles in the V0 detector.

@@

%The multiplicity class is classified by the signal amplitude of the V0 detector.
The high multiplicity events used in the present analysis have the top 0.1\% signal amplitude (the 0--0.1\% multiplicity class) by the V0 detector with respect to the set of minimum bias events at ALICE. Corresponding number of charged particles is reported as $\sim$31 in the mid-rapidity region ($|\eta|<0.5$) that is almost 5 times larger than the minimum-bias events. Statistics of the high multiplicity event benefits from a dedicated high multiplicity trigger that was recently implemented at ALICE  in the LHC Run 2 period.

@@

The two-particle correlations function between the trigger and associated particles are measured as functions of relative pseudorapidity ($\Delta\eta$) and azimuthal angle differences($\Delta\varphi$)
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d} \Delta\eta \rm{d}\Delta\varphi} = B(0, 0)\frac{S(\Delta\eta, \Delta\varphi)}{B(\Delta\eta, \Delta\varphi)}\quad,
\label{eq:corrfunction}
\end{eqnarray}
where  $\pttrig$ and $\ptassoc$ are the transverse momenta of  the trigger and associated particles, respectively, $N_\mathrm{trig}$ is the number of trigger particles, $N_\mathrm{pair}$ is the number of pairs of the trigger and associated particles, and $S (\Delta\eta, \Delta\varphi)$ and $B (\Delta\eta, \Delta\varphi)$ are the signal and background distributions, respectively. $S (\Delta\eta, \Delta\varphi)$ is constructed using two-particle correlations in the same event and $B(\Delta\eta, \Delta\varphi)$ is in mixed events. The background distribution is normalized by its value at $\Delta\eta$ and $\Delta\varphi = 0$, represented as $B (0,0)$. In mixed events, the primary vertices and multiplicity classes are required to be in the same range to avoid the acceptance effect with respect to the multiplicity class and primary vertex.

The ridge yields at large $\Delta\eta$ are extracted in different multiplicity classes and transverse momentum intervals. The large $\Delta\eta$ range is selected as $1.6<|\Delta\eta|<1.8$ where non-flow effects (mainly coming from jets) are negligible. In this region, the $\Delta\varphi$ distribution, or the so called per-trigger yield is expressed as
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } = \int_{|\Delta \eta|>1.6} \rm{d} \Delta \eta \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d}\Delta\eta d\Delta\varphi} \quad.
\end{eqnarray}

The baseline of the correlations is subtracted by implementing Zero-Yield-At-Minimum (ZYAM) procedure~\cite{Ajitanand:2005jj}. The minimum yield $(C_{\rm{ZYAM}})$ at the $\Delta\varphi$ ($\Delta\varphi_{\rm{min}}$) in the $\Delta\varphi$ distribution is obtained from the function, which fits the $\Delta\varphi$ distribution with Fourier series up to the third harmonic. Subtracting $C_{\rm{ZYAM}}$ from the $\Delta\varphi$ distribution makes the yield at $\Delta\varphi_{\rm{min}}$ zero in order to describe the shape to find the minimum to be used for integral of the yield. The ridge yield ($Y^{\rm{assoc}}$) is obtained by integrating the near-side peak of the $\Delta\varphi$ distribution over $|\Delta\varphi|<|\Delta\varphi_{\rm{min}}|$ after the ZYAM procedure
\begin{eqnarray}
Y^{\rm{assoc}} = \int_{|\Delta \varphi| < |\Delta\varphi_{\rm{min}}| } \rm{d} \Delta\varphi \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } \quad.
\end{eqnarray}

The ridge yield is further studied by exploiting the largest momentum transfer among the initial partons in a given event, which results from the hard scattering. In this article, larger momentum transfer denotes harder event scale. The larger momentum transfer is also expected to be connected with the shorter impact parameters in pp collisions in average~\cite{Sjostrand:1986ep,Frankfurt:2010ea} so that the event scale can be indirectly connected with the impact parameters.
% The ridge yield is further studied by exploiting the momentum transfer between the interacting partons,so called, event scale.  On average, increasingly hard parton interactions result from pp collisions with decreasing impact parameters between the two proton ///original version
The event scales are set by requiring minimum transverse momentum of leading track or reconstructed jet in mid-rapidity since hard scattering produces leading tracks with a high $p_\mathrm{T}$ or jets. The leading track is selected within $|\eta|<0.9$ and the full azimuthal angle. Jets are reconstructed with charged particles only (charged jets) for the anti-$k_{\rm{T}}$ algorithm and the resolution parameter $R = 0.4$. Jets are selected in $|\eta_\mathrm{jet}|<0.4$ and the full azimuthal angle. The transverse momentum of jets $p_\mathrm{T,\,jet}$ is corrected for underlying event density that is measured for the $k_{\rm{T}}$ algorithm with $R = 0.2$.

@@

\section{Systematic Uncertainty}
\label{sec:uncertainties}
The systematic uncertainty is evaluated by varying the event selection, track selection, yield extraction procedures and detector efficiency correction. Each source is assumed to be uncorrelated. The uncertainty from the event selection can be divided into pile-up rejection and primary vertex selection. The uncertainty from the track selection is estimated by changing the selection criteria. The uncertainty from the yield extraction procedures can be categorized into long-range definition and ZYAM procedure. The uncertainty from the efficiency correction is estimated by enlarging the event mixing bins, which is relevant to the acceptance correction.
%Finally, The uncertainty from the application of the detector efficiency correction is estimated comparing the results constructed by efficiency corrected tracks with the results constructed. 
Finally, the uncertainty from the efficiency correction for charged tracks is estimated by comparing the correlations function constructed by reconstructed tracks from the detector responses with the correlation function constructed by generated particles from the event generator.
The uncertainty of $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$ is summarized by averaging each uncertainty for given $\pttrig$ and $\ptassoc$. The uncertainty of $Y^{\rm{assoc}}(\ptlead)$ and $Y^{\rm{assoc}}(\ptjet)$ is summarized by averaging the each uncertainty for a given $\ptlead$ and $\ptjet$ selection, respectively.
\begin{table}[h!]
\caption{The relative systematic uncertainty of the associated yield spectrum are estimated for $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$, $Y^{\rm{assoc}}(\ptlead)$, and $Y^{\rm{assoc}}(\ptjet)$, respectively.}
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
\multirow{2}{*}{Sources}  & \multicolumn{3}{c|}{Systematic uncertainty (\%)} \\\cline{2-4}
         & $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ \\ \hline \hline
Pileup rejection		& $\sim$0.8	&0--4		&0--4		\\ \hline
Primary Vertex		& $\sim$2.4	&1--12	&1-8		\\ \hline

@@

The uncertainty from the ZYAM procedure is estimated by varying the finding-scope for minimum position. The estimated uncertainty for track selection is 5.1\% by averaging the uncertainty for the whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks, which is not largely affected by the event scale selection. The uncertainty from the $\Delta\eta$ projection range is estimated by varying the $\Delta\eta$ range. The uncertainty might include effects from the longitudinal de-correlations and possible non-flow contamination, which is thought be small. The estimated uncertainty is increased up to 9\% from 3\%, which is obtained by averaging the whole transverse momentum range. The uncertainty from the mixing pool bin size of primary vertex is estimated by varying the bin size from 2 to 1 cm. The estimated uncertainty is increased up to 8-16\% from 1-2\%.

>>>>

%LHC at CERN produces various interesting phenomena including the ridge effect that is the main topic of the document in pp collisions with the highest center-of-mass energy in the world.
%Recent center-of-mass energy in pp collisions reaches up to $\sqrt{s} = 13$~TeV during the last LHC Run 2 period. 
The analysis is based on the data sets of pp collisions at $\sqrt{s} = 13$~TeV collected from 2016 to 2018 during the LHC Run 2 period. The full description of the ALICE detector in the LHC Run 2 can be found in Refs.~\cite{Aamodt:2008zz,Abelev:2014ffa}. The present analysis mainly utilizes the V0~\cite{Abbas:2013taa}, ITS (Inner Tracking System)~\cite{aliceITS}, and TPC (Time Projection Chamber)~\cite{aliceTPC} detectors.

The V0 detector consists of two rings, V0-A and V0-C, each made of 32 scintillator tiles, covering the full azimuthal angle within pseudorapidity intervals $2.8 < \eta < 5.1$ and $-3.7 < \eta < -1.7$, respectively. The V0 provides trigger and estimation of event multiplicity. A sample of events including higher numbers of produced particles is obtained with a high-multiplicity trigger, which is achieved by requiring 5 times of mean signal amplitudes of energy deposit by charged particles in the V0 detector.

@@

%The multiplicity class is classified by the signal amplitude of the V0 detector.
The high-multiplicity events used in the present analysis have the top 0.1\% signal amplitude (the 0--0.1\% multiplicity class) by the V0 detector with respect to the set of minimum bias events at ALICE. Corresponding number of charged particles is reported as $\sim$31 in the mid-rapidity region ($|\eta|<0.5$) that is almost 5 times larger than the minimum-bias events. Statistics of the high-multiplicity event benefits from a dedicated high-multiplicity trigger that was recently implemented at ALICE  in the LHC Run 2 period.

@@

The two-particle correlations function is measured as functions of relative pseudorapidity ($\Delta\eta$) and azimuthal angle differences($\Delta\varphi$) between the trigger and associated particles.
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d} \Delta\eta \rm{d}\Delta\varphi} = B(0, 0)\frac{S(\Delta\eta, \Delta\varphi)}{B(\Delta\eta, \Delta\varphi)} \lvert_{\pttrig,\ptassoc}  , 
\label{eq:corrfunction}
\end{eqnarray}
where  $\pttrig$ and $\ptassoc$ are transverse momenta of the trigger and associated particles, respectively, $N_\mathrm{trig}$ is the number of trigger particles, $N_\mathrm{pair}$ is the number of trigger-associated particle pairs, and $S (\Delta\eta, \Delta\varphi)$ and $B (\Delta\eta, \Delta\varphi)$ are the signal and background distributions, respectively. $S (\Delta\eta, \Delta\varphi)$ is constructed using two-particle correlations in the same event and $B(\Delta\eta, \Delta\varphi)$ is in mixed events. The background distribution is normalized by its value at $\Delta\eta$ and $\Delta\varphi = 0$, represented as $B (0,0)$. In mixed events, the primary vertices and multiplicity classes are required to be in the same range to avoid the acceptance effect with respect to the multiplicity class and primary vertex.

The ridge yields at large $\Delta\eta$ are extracted in different multiplicity classes and transverse momentum intervals. The large $\Delta\eta$ range is selected as $1.6<|\Delta\eta|<1.8$ where non-flow effects (mainly coming from jets) are negligible. In this region, the $\Delta\varphi$ distribution, or the so called per-trigger yield is expressed as
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } = \int_{1.6<|\Delta \eta|<1.8} \rm{d} \Delta \eta \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d}\Delta\eta d\Delta\varphi} \quad.
\end{eqnarray}

The baseline of the correlations is subtracted by implementing Zero-Yield-At-Minimum (ZYAM) procedure~\cite{Ajitanand:2005jj}. The minimum yield $(C_{\rm{ZYAM}})$ at the $\Delta\varphi$=$\Delta\varphi_{\rm{min}}$ in the $\Delta\varphi$ distribution is obtained from the function, which fits the $\Delta\varphi$ distribution with a Fourier series up to the third harmonic. Subtracting $C_{\rm{ZYAM}}$ from the $\Delta\varphi$ distribution makes the yield at $\Delta\varphi_{\rm{min}}$ zero. The ridge yield ($Y^{\rm{assoc}}$) is obtained by integrating the near-side peak of the $\Delta\varphi$ distribution over $|\Delta\varphi|<|\Delta\varphi_{\rm{min}}|$ after the ZYAM procedure
\begin{eqnarray}
Y^{\rm{assoc}} = \int_{|\Delta \varphi| < |\Delta\varphi_{\rm{min}}| } \rm{d} \Delta\varphi \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } \quad.
\end{eqnarray}

The ridge yield is further studied in events having a hard jet or a high-pT leading particle in the ALICE central barrel. Such requirement is expected to bias impact parameter of pp collisions to be smaller on average~\cite{Sjostrand:1986ep,Frankfurt:2010ea}.
%The ridge yield is further studied by exploiting the largest momentum transfer among the initial partons in a given event, which results from the hard scattering. In this article, larger momentum transfer denotes harder event-scale. The larger momentum transfer is also expected to be connected with the shorter impact parameters in pp collisions in average~\cite{Sjostrand:1986ep,Frankfurt:2010ea} so that the event-scale can be indirectly connected with the impact parameters.
% The ridge yield is further studied by exploiting the momentum transfer between the interacting partons,so called, event scale.  On average, increasingly hard parton interactions result from pp collisions with decreasing impact parameters between the two proton ///original version
The event-scale is set by requiring minimum transverse momentum of leading track or reconstructed jet in mid-rapidity since hard scattering produces leading tracks with a high $p_\mathrm{T}$ or jets. The leading track is selected within $|\eta|<0.9$ and the full azimuthal angle. Jets are reconstructed with charged particles only (charged jets) with the anti-$k_{\rm{T}}$ algorithm and the resolution parameter $R = 0.4$. Jets are selected in $|\eta_\mathrm{jet}|<0.4$ and the full azimuthal angle. The transverse momentum of jets $p_\mathrm{T,\,jet}$ is corrected for underlying event density that is measured for the $k_{\rm{T}}$ algorithm with $R = 0.2$.

@@

\section{Systematic Uncertainties of the Measured Yields}
\label{sec:uncertainties}
Systematic uncertainties were estimated by varying the event selection, track selection, yield extraction procedures and detector efficiency correction. Each source is assumed to be uncorrelated. The uncertainty from the event selection can be divided into pile-up rejection and primary vertex selection. The uncertainty from the track selection is estimated by changing the selection criteria. The uncertainty from the yield extraction procedures can be categorized into long-range definition and ZYAM procedure. The uncertainty from the efficiency correction is estimated by enlarging the event mixing bins, which is relevant to the acceptance correction.
%Finally, The uncertainty from the application of the detector efficiency correction is estimated comparing the results constructed by efficiency corrected tracks with the results constructed. 
Finally, the uncertainty from the efficiency correction for charged tracks is estimated by comparing the correlations function constructed by reconstructed tracks from the detector responses with the correlation function constructed by generated particles from the event generator.
The uncertainty of $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$ is summarized by averaging each uncertainty for given $\pttrig$ and $\ptassoc$. The uncertainty of $Y^{\rm{assoc}}(\ptlead)$ and $Y^{\rm{assoc}}(\ptjet)$ is summarized by averaging the each uncertainty for a given $\ptlead$ and $\ptjet$ selection, respectively.
\begin{table}[h!]
\caption{The relative systematic uncertainty of the associated yield spectrum are estimated for $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$, $Y^{\rm{assoc}}(\ptlead)$, and $Y^{\rm{assoc}}(\ptjet)$, respectively.}
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
\multirow{2}{*}{Sources}  & \multicolumn{3}{c|}{Systematic uncertainty (\%)} \\\cline{2-4}
         & $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ \\ \hline \hline
Pileup rejection		& $\sim$0.8	&0--4		&0--4		\\ \hline
Primary Vertex		& $\sim$2.4	&1--12	&1-8		\\ \hline

@@

The uncertainty from the ZYAM procedure is estimated by varying the finding-scope for minimum position. The estimated uncertainty for track selection is 5.1\% by averaging the uncertainty for the whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks, which is not largely affected by the event-scale selection. The uncertainty from the $\Delta\eta$ projection range is estimated by varying the $\Delta\eta$ range. The uncertainty might include effects from the longitudinal de-correlations and possible non-flow contamination, which is thought be small. The estimated uncertainty is increased up to 9\% from 3\%, which is obtained by averaging the whole transverse momentum range. The uncertainty from the mixing pool bin size of primary vertex is estimated by varying the bin size from 2 to 1 cm. The estimated uncertainty is increased up to 8-16\% from 1-2\%.
